---
title: "Business Intelligence Lab Submission Markdown"
author: "<Specify your name here>"
date: "<Specify the date when you submitted the lab>"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: inline
---

# Student Details

|                                                   |                                             |
|---------------------------|---------------------------------------------|
| **Student ID Numbers and Names of Group Members** |  134879 Tulienge Lesley                     |
| **BBIT 4.2 Group**                                |  Group C                                    |
| **Course Code**                                   | BBT4206                                     |
| **Course Name**                                   | Business Intelligence II                    |
| **Program**                                       | Bachelor of Business Information Technology |
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023  |

# Setup Chunk

**Note:** the following KnitR options have been set as the global defaults:\
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE,                        collapse = FALSE, tidy = TRUE)`.

> output:\
> \
> github_document:\
> toc: yes\
> toc_depth: 4\
> fig_width: 6\
> fig_height: 4\
> df_print: default\
> \
> editor_options:\
> chunk_output_type: console

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE,
                      collapse = FALSE, tidy = TRUE)
```

# Load the Datasets

This code chunk loads a dataset with 101 rows and 100 columns using the readr library, displaying it in RStudio with column specifications. This data is well-structured, with numeric and categorical variables. It contains ratings, scores, and feedback related to different aspects of the course and students' performance. The data also includes some missing values indicated as "NA's."

```{r Your Second Code Chunk, echo=FALSE}
library(readr)
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset <- read_csv("data/20230412-20230719-BI1-BBIT4-1-StudentPerformanceDataset.csv")
View(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)
```

#### Preview the Loaded Datasets

This code chunk uses the **dim** function from the readr library to display the dimensions of the dataset **X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset**. It shows the number of rows (observations) and columns (attributes/variables/features) in the dataset.

```{r Your Third Code Chunk}
library(readr)
dim(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)

```

#### Identify the Data Types

This code chunk uses the **sapply** function from the readr library to determine and display the data types (classes) of each column in the dataset **X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset**. Understanding data types is crucial for selecting appropriate visualization techniques and data analysis methods, especially when dealing with factors or numeric data that may require specific handling or conversion.

```{r Your Fourth Code Chunk, eval=FALSE, include=FALSE}
library(readr)
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset, class)
```

# Descriptive Statistics

Descriptive statistics are a set of techniques and metrics used to summarize and describe the main features, characteristics, and patterns within a dataset. These statistics provide a snapshot of the data's distribution, central tendency, variability, and other important properties.Before you can effectively use your data for tasks like building prediction models or drawing meaningful conclusions, it's crucial to have a deep understanding of the dataset.Understanding the context is essential for making informed decisions based on those results

#### Measures of Frequency

This code calculates and displays the frequency and mode (most common value) for several categorical variables in a dataset, helping to understand the data's distribution and identify prevalent values.Some datasets show relatively even distributions, while others have dominant categories. The frequencies provide insights into the composition of each dataset.

```{r 5th}
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq <- X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$class_group
cbind(frequency = table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq),
      percentage = prop.table(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq)) *100)

X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq <- X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$alcohol_or_narcotics
cbind(frequency = table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq),
      percentage = prop.table(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq)) *100)

X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq <- X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$financial_wellness
cbind(frequency = table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq),
      percentage = prop.table(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq)) *100)

X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq <- X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$romantic_relationships
cbind(frequency = table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq),
      percentage = prop.table(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq)) *100)

X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq <- X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$gender
cbind(frequency = table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq),
      percentage = prop.table(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_freq)) *100)

X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_class_group_mode <- names(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$class_group))[
  which(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$class_group) == max(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$class_group)))
]
print(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_class_group_mode)

X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_alcohol_or_narcotics_mode <- names(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$alcohol_or_narcotics))[
  which(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$alcohol_or_narcotics) == max(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$alcohol_or_narcotics)))
]
print(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_alcohol_or_narcotics_mode)

X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_financial_wellness_mode <- names(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$financial_wellness))[
  which(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$financial_wellness) == max(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$financial_wellness)))
]

print(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_financial_wellness_mode)

X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_romantic_relationships_mode <- names(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$romantic_relationships))[
  which(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$romantic_relationships) == max(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$romantic_relationships)))
]
print(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_romantic_relationships_mode)

X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_gender_mode <- names(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$gender))[
  which(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$gender) == max(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$gender)))
]
print(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_gender_mode)

```
# Measures of Central Tendency
### - Calculate the mode of studying in study group

```{r code}
studentperformance_mode <- names(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$studying_in_study_group))[
  which(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$studying_in_study_group) == max(table(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset$studying_in_study_group)))
]
print(studentperformance_mode)

```

#### Measures of distribution

The summary function provides key statistics and information for each variable, such as mean, median, quartiles, and count, helping to assess the distribution and characteristics of the data

```{r 6th code}
summary(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)
```

#### Measures of standard deviation

This code calculates and returns the standard deviation for each of the selected columns (variables) in the dataset. Standard deviation measures the dispersion or spread of data points around the mean, providing insight into data variability

```{r 7th code}
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset[, c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99)], sd)

```

#### Measures of variance

This code calculates and returns the variance for each selected column. Variance quantifies how much individual data points deviate from the mean, offering another perspective on data dispersion.

```{r 8th code}
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset[, c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99)], var)
```

#### Measures of kurtosis

This code calculates the kurtosis for the selected columns, which indicates the degree of data distribution's peakedness or tail heaviness.The Kurtosis informs you of how often outliers occur in the results

```{r 9th code}
if (!is.element("e1071", installed.packages()[, 1])) {
  install.packages("e1071", dependencies = TRUE)
}
require("e1071")
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset [c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99)],kurtosis, type=2)
```

#### Measures of skewness

This code calculates the skewness for the selected columns, indicating the asymmetry in the data distribution. Positive skewness indicates a right-skewed distribution, while negative skewness suggests a left-skewed distribution 1. Skewness between -0.4 and 0.4 (inclusive) implies that there is no skew in the distribution of results; the distribution of results is symmetrical; it is a normal distribution. 2. Skewness above 0.4 implies a positive skew; a right-skewed distribution. 3. Skewness below -0.4 implies a negative skew; a left-skewed distribution

```{r 10th code}
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset [c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99)],skewness, type=2)
```

# Measures of Relationship

#### Measures of covariance

This code calculates the covariance matrix, which measures the degree of linear association or dependence between pairs of selected variables. The resulting covariance matrix is then displayed

```{r 11th code}
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_cov <- cov(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset [c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99)])
View(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_cov)
```

#### Measures of correlation

This code calculates the correlation matrix, which measures the strength and direction of linear relationships between pairs of selected variables. The resulting correlation matrix is then displayed.

```{r 12th code}
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_cor <- cor(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset [c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44, 45, 46, 47, 48, 49, 50, 51, 52, 53, 54, 55, 56, 57, 58, 59, 60, 61, 62, 63, 64, 65, 66, 67, 68, 69, 70, 71, 72, 75, 76, 77, 78, 79, 80, 81, 82, 83, 84, 85, 86, 87, 88, 89, 90, 91, 92, 93, 94, 95, 96, 97, 98, 99)])
View(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_cor)
```

# Inferential Statistics

```{r 13th code}


```
